\newpage
\section{Dimensional Splitting}


To go from one to multiple dimensions, it is tempting to just extend the one dimensional discretisation.
For example, starting with the conservation law

\begin{equation}
	\deldt{\U} + \deldx{\F(\U)} + \frac{\del}{\del y} \mathbf{G}(\U) = 0
\end{equation}


and just apply Godunov's finite volume method:

\begin{equation}
	\U_{i,j}^{n+1} = 
					\U_{i,j}^n 
					+ \frac{\Delta t}{\Delta x} \left( \F_{i - \half, j} - \F_{i + \half, j}\right) 
					+ \frac{\Delta t}{\Delta x} \left( \mathbf{G}_{i ,j - \half} - \mathbf{G}_{i, j + \half}\right) 
\end{equation}


However, this is a bit of a problem.
The upwinding here is not complete.
Consider the 2D advection equation

\begin{equation}
	\deldt{q} + u\deldx{q} + v \frac{\del}{\del y} q = 0
\end{equation}

Now suppose we have advecting velocities $u = v = 1$, i.e. the advection velocity is along the diagonal.
Then our method reads 

\begin{equation}
	q_{i,j}^{n+1} = 
					q_{i,j}^n 
					+ u \frac{\Delta t}{\Delta x} \left( q_{i - \half, j} - q_{i + \half, j}\right) 
					+ v \frac{\Delta t}{\Delta x} \left( q_{i ,j - \half} - q_{i, j + \half}\right) 
\end{equation}

This expression doesn't involve $q_{i-1, j-1}$ at all, but that's actually the value that should be advected to $q_{i, j}$ in the next timestep!


One way of doing things is to actually formulate more sophisticated unsplit methods that involve the appropriate stencils.
We shan't do that here though.
We make use of dimensional splitting. 
Instead of solving

\begin{equation}
	\begin{cases}
		\text{PDE: } & \deldt{\U} + \deldx{\F(\U)} + \frac{\del}{\del y} \mathbf{G}(\U) = 0\\
		\text{IC: } & \U(x, y, t^n) = \U^n
	\end{cases}
\end{equation}

we do it in 2 (3 for 3D) steps:

Step 1: We obtain a intermediate result $\U^{n+\half}$ by solving the ``x - sweep'' over the full time interval $\Delta t$ :
\begin{equation}
	\begin{cases}
		\text{PDE: } & \deldt{\U} + \deldx{\F(\U)} = 0\\
		\text{IC: } &  \U^n
	\end{cases}
\end{equation}

and then we evolve the solution to the final $\U^{n+1}$ by solving the ``y - sweep'' over the full time interval $\Delta t$ :

\begin{equation}
	\begin{cases}
		\text{PDE: } & \deldt{\U} + \frac{\del}{\del y} \mathbf{G}(\U) = 0\\
		\text{IC: } & \U^{n+\half} \\
	\end{cases}
\end{equation}

using the 1D methods that are described.

What's even better is that it can be shown (see \cite{leveque_2002}) that by switching the order of the sweeps every timestep (and keeping the time step interval $\Delta t$ constant) leads to a second order accurate method in time.
This is called ``Strang splitting''.
















